A new tidy data structure to support exploration and modeling of temporal data

### 2019-08-12 tsibble.Rmd
done with revision on subsection "key"

Figure \@ref(fig:tsibble-semantics) sketches out the data form required for a tsibble, an equivalent of the "tidy" format. Beyond the layout, tsibble gives the contextual meaning to variables in order to construct the temporal data object, as newly introduced "index" and "key" semantics stated in the first and second definitions above. Variables other than "index" and "key" are considered as measurements. The third and fourth definitions imply that tsibble is *tidier* than "tidy data", placing itself as a model input that gives rise to more robust and reliable downstream analytics.

@@

The "key" specification is the second essential ingredient for a tsibble. The "key" uniquely identifies observations that are recorded over time in a data table. It is similar to a primary key [@codd_relational_1970] defining each observation in a relational database. In the wider format in which multiple time series are often structured, the columns hold a series of values, so that the column implicitly serves as identification. In longer format, columns are melted with names converted to "key" values. However, the "key" provides much more flexibility. It is not constrained to a single field, but can be composed from multiple fields. The identifying variables from which the "key" is constituted remain the same as in the original table with no further tweaks.

The "key" is usually known a priori by analysts. For example, Table \@ref(tab:tb-sub) describes the number of tuberculosis cases for each gender across the countries every year. This data description suggests that columns `gender` and `country` have to be declared as the key, similar to a panel variable for longitudinal data. Lacking either of the two will be inadequate, because the observations would not be uniquely identified, and thus a tsibble construction would fail. An alternative specification of the key for this data is to include a third variable `continent`. Since `country` is nested within `continent`, it is a free variable for use. This variable brings additional information that can be used for forecasting reconciliation [@fpp]. The key needs to be explicit when multiple units exist in the data. The key can be implicit when it finds a univariate series in the table, but it cannot be absent from a tsibble.

The "key" also provides a link between the data, models, and forecasts. This neatly decouples the data from models and forecasts, leaving more room for necessary model components, such as coefficients, fitted values and residuals. More details are given  in Section \@ref(sec:ts-models).

@@

>>>>

Figure \@ref(fig:tsibble-semantics) sketches out the data form required for a tsibble, an equivalent of the "tidy" format. Beyond the layout, tsibble gives the contextual meaning to variables in order to construct the temporal data object, as newly introduced "index" and "key" semantics stated in the definitions 1 and 2 above. Variables other than "index" and "key" are considered as measurements. The definitions of 1 and 2 imply that tsibble is *tidier* than "tidy data", positioning itself as a model input that gives rise to more robust and reliable downstream analytics.

@@

> *Key is a set of variables that define observational units over time.*

What subjects/entities to be observed over time leads to the second component of a tsibble--key. Key can consist of empty, single, or multiple variables identifying units measured along the way. When only a single observational unit is present in the table, key can be left with no variables to be specified. But when multiple units exist in the data, key should be supplied by identifying variables to sufficiently define the units. In longitudinal data, key can be thought of as "panel" (such as in the Stata) constrained to a single variable. In tsibble, key allows for multiple variables of nesting and crossing relations [@wilkinson2006grammar] to be useful for forecasting reconciliation [@fpp; @R-hts] and other advanced models. For example, Table \@ref(tab:tb-sub) describes the number of tuberculosis cases for each gender across the countries every year. This suggests that the key comprises at least columns `gender` and `country`. Since `country` is nested within `continent`, it is a free variable to be included in the key specification.

> *Each observation should be uniquely identified by index and key.*

Inspired by "primary key" [@codd_relational_1970], a unique identifier for each observation in a relational database, tsibble key also uniquely identifies each observational unit over time. Any duplicates of key-index pairs will fail upon the construction of a tsibble. This is, either `gender` or `country` alone is not enough to be the key for the tuberculosis data. The rigidity of tsibble, as the fundamental data infrastructure, warrants the validity of temporal data analysis for the later stage. It highlights the primary difference from *time tibble* provided by **tibbletime**.

Since observational units are embedded, modelling and forecasting for each unit in a tsibble will be accelerated. The tsibble key plays a role of the central transit hub in connecting multiple tables managed by the data, models, and forecasts. This neatly decouples the data copy from models and forecasts, significantly reducing the storage space. More details are given in Section \@ref(sec:ts-models).

@@

> *Each observational unit should be measured at a common interval, if regularly spaced.*
